---
title: "DSTI_TimeSeries"
output: html_notebook
---


```{r}
install.packages("readxl")
install.packages("fpp2")
```

```{r}
library(ggfortify)
library(readr)
library(forecast)
```


```{r}
'data_raw <- read_excel("data\\raw\\Elec-train.xlsx")
data <- read_excel("data\\processed\\Elec-train-processed-2.xlsx", col_names = TRUE, col_types = c("date", "numeric", "numeric"))
'
data <- read_csv2("data\\processed\\Elec-train-processed-2.csv", col_names = TRUE, col_types = c("n", "n"))
```

```{r}
autoplot(as.ts(data))
```



```{r}
# convert date column to date class
data$Timestamp <- as.POSIXlt(data$Timestamp, format = "%m/%d/%y  %H:%M")

# view R class of data
class(data$Timestamp)
## [1] "Date"

# view results
head(data$Timestamp)
```


```{r}
data_known <- ts(data[1:4603,], frequency = 96)
data_to_forecast <- ts(data[4604:4699,], frequency = 96)
#data_known_5 <- ts(data[1:4603,], frequency = 5)

#$Power <- ts(data$Power, frequency = 96)
#data$Temp <- ts(data$Temp, frequency = 96)
#Power <- data$Power
#Temp <- data$Temp'
```


First of all we ensure that the TimeStamp column has always the same duration between all TimeStamp.

```{r}
plot(data$Power,type='l',xlim=c(1,4603),ylim=c(120,360),xlab='Time',ylab='Power (kW)')
```


```{r}
plot(data$Temp,type='l',xlim=c(1,4603),ylim=c(4,20),xlab='Time',ylab='Temperature (°C)')
```

```{r}
autoplot(data)
```

Timestamp per day = 96
Nb total rows for Power = 4603 => 47 days + 91 timestamps
Train => 80% => 3682.4 => floor(3682.4) => 3682 => 38 days + 34 timestamps
Test => 20% => 920.6 => floor(920.6) => 921 => 9 days + 57 timestamps


```{r}
elec_train=window(data_known,start=c(1,1),end=c(39,35))
elec_test=window(data_known,start=c(39,36),end=c(48,91))
```

Other definition

```{r}

h2 <- 921L
#h2 <- 184L
elec_train2 <- head(data_known, round(length(data_known)/2 - h2))
elec_test2 <- tail(data_known, h2)
```

```{r}
elec_train <- elec_train2
elec_test <- elec_test2
```


```{r}
data_known_Power <- data_known[,1]
elec_train_Power <- elec_train[,1]
elec_test_Power <- elec_test[,1]
data_known_Temp <- data_known[,2]
elec_train_Temp <- elec_train[,2]
elec_test_Temp <- elec_test[,2]
elec_forcast_Temp <- data_to_forecast[,2]
summary(data_known_Power)
summary(data_known_Temp)
```
```{r}
ggtsdisplay(data_known_Power)
```
Seasonal pattern

```{r}
plot(data_known_Power[1:96],type="l",ylim=c(min(data_known_Power),max(data_known_Power)))
for (i in 1:4603) lines(data_known_Power[(1+96*i):(96*(i+1))])
```




FORECASTING WITHOUT COVARIATES (Power)

1- Holt-Winters exponential smoothing

```{r}
fit1=hw(elec_train_Power,lambda="auto")
prev1=forecast(fit1,h=921)
autoplot(prev1) + autolayer(elec_test_Power, series="true data")+
autolayer(prev1$mean, series="HW forecasts")
```
Frequency is too high to use Holt-Winters = Holt-Winters is really made for shorter seasonal periods.

2- SARIMA model

```{r}
fit2=auto.arima(elec_train_Power,lambda="auto")
prev2=forecast(fit2, h=921)
autoplot(elec_test_Power)+autolayer(prev2$mean,series="SARIMA without covariate")
```
RMSE :
```{r}
print(sqrt(mean((prev2$mean-elec_test_Power)^2)))
```


```{r}
autoplot(diff(elec_train_Power,lag = 96))
```


Forecasting with covariates

3 - Auto Arima

```{r}
elec_train_Temp_ts <- elec_train[,1] <- list(NULL) 
elec_train_Temp_ts
```


```{r}
fit3=auto.arima(elec_train[,"Power"],xreg=elec_train[,"Temp"])
prev3=forecast(fit3,h=921,xreg=elec_test_Temp)
autoplot(elec_test_Power)+autolayer(prev3$mean)
```
```{r}
print(sqrt(mean((prev3$mean-elec_test_Power)^2)))
```
```{r}
summary(fit3)
```
```{r}
checkresiduals(fit3)
```
P-value is very low and it seems that there is not really a relationship between Power and Temp :
```{r}
plot(elec_train[,"Power"],elec_train[,"Temp"])
```

## Forecasting with a auto-regressive neural network

```{r}
fit4=nnetar(elec_train[,"Power"],T=96)
print(fit4)
prevNN=forecast(fit4,h=921)
autoplot(elec_test_Power)+autolayer(prevNN$mean)
```
RMSE :
```{r}
print(sqrt(mean((prevNN$mean-elec_test_Power)^2)))
```

Manual improvements as the result proposed by the default settings are not suitable :

```{r}
fit5=nnetar(elec_train[,"Power"],p=8,P=10,T=96)
print(fit5)
prevNN5=forecast(fit5,h=921)
autoplot(elec_test_Power)+autolayer(prevNN5$mean)
```

RMSE :
```{r}
print(sqrt(mean((prevNN5$mean-elec_test_Power)^2)))
```

Forecast based on Power only with NN

```{r}
autoplot(forecast(fit5,h=96))
```

## Forecasting with a auto-regressive neural network with covariates

```{r}
fit5=nnetar(elec_train[,"Power"], xreg=elec_train[,"Temp"],  T=96)
print(fit5)
prevNN5=forecast(fit5, xreg=elec_test[,"Temp"], h=921)
autoplot(elec_test_Power)+autolayer(prevNN5$mean)
```

RMSE :
```{r}
print(sqrt(mean((prevNN5$mean-elec_test_Power)^2)))
```

Manual improvements as the result proposed by the default settings are not suitable :

```{r}
fit6=nnetar(elec_train[,"Power"], xreg=elec_train[,"Temp"],p=8,P=10,T=96)
print(fit6)
prevNN6=forecast(fit6, xreg=elec_test[,"Temp"],h=921)
autoplot(elec_test_Power)+autolayer(prevNN6$mean)
```

RMSE :
```{r}
print(sqrt(mean((prevNN6$mean-elec_test_Power)^2)))
```

Forecast based on Power and Temp with NN

```{r}
forecast_NN_Cov <- forecast(fit6,xreg=elec_forcast_Temp,h=96)
autoplot(forecast_NN_Cov)
```
